


\chapter*{ABSTRACT}
\addcontentsline{toc}{chapter}{ABSTRACT}




\begin{SingleSpace}
	


\initial{I}n this study, we examined the factors associated with weight gain among people living with HIV receiving dual therapy, as part of the MODERATO trial, the first therapeutic simplification trial conducted in Africa. The study, carried out in Côte d’Ivoire, Burkina Faso, and Cameroon, included 320 patients randomly assigned to two groups: Dolutegravir + Lamivudine (DTG+3TC) and Atazanavir/ritonavir + Lamivudine (ATV/r+3TC), and followed over a period of 96 weeks. \vspace{0.2cm}

To address this issue, we applied two statistical approaches. A linear mixed model was used to analyze weight changes over time, while a Cox proportional hazards regression model was implemented to identify factors associated with a clinical event defined as a weight gain of at least 5\% from baseline. \vspace{0.2cm}

The study results showed that the Dolutegravir-based regimen (DTG+3TC) was associated with a significantly greater weight gain than the Atazanavir/ritonavir-based regimen (ATV/r+3TC). The rate of weight gain was estimated at 553 grams per 100 days under DTG+3TC, compared to 360 grams per 100 days under ATV/r+3TC, corresponding to an annual difference of approximately 704.45 grams. The multivariate Cox model analysis identified three main risk factors for rapid weight gain ($\geqslant 5\%$). First, female sex, with a 52\% higher risk compared to men. Second, low baseline weight, as each additional kilogram reduced the risk by 2\%, supporting the “return to health” hypothesis among initially underweight patients. Finally, geographical location: patients followed in Burkina Faso had a 40\% higher risk than those followed in Côte d’Ivoire, suggesting the influence of local factors.\vspace{0.2cm}

These findings highlight the importance of a personalized medicine approach, with enhanced metabolic monitoring for at-risk populations, particularly women and patients with low baseline weight, especially when treated with DTG. Such a proactive strategy is essential to prevent long-term cardiometabolic complications while maintaining the excellent virological outcomes of these simplified therapeutic regimens. \vspace{0.5cm}

\textbf{Keywords}: HIV, dual therapy, dolutegravir, atazanavir, weight gain, West Africa.


\end{SingleSpace}
\clearemptydoublepage

